---
title: "R Notebook"
output: html_notebook
---

# Read Me
This script models five scenarios.
1) Aggregate utilitarian: everyone grows the same
2) Prioritarian: lowest 5 ones grow while other stay constant; the lower the higher growth
distances between lowest 5 curves must decline; no guarantee for reaching threshold
3) Egalitarian: convergence to one point
4) sufficientarian: lower ones grow to threshold by 2040 and stay there, highest stay constant
5) Limitarian: highest ones reduce, lowers stay constant
When core logic of justice principle doesn't specify trajectory for some regions, regions stay constant

```{r load libraries and set up environment}
#import libraries
library(reticulate)
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)

require(reticulate)
#IIASA
use_condaenv(condaenv = "C:\\Users\\scheifinger\\AppData\\Local\\anaconda3\\envs\\Streamlit", conda = "auto", required = NULL)
#home
#use_condaenv("C:\\Users\\schei\\anaconda3\\envs\\streamlit", conda="auto", required = NULL)
pyam <- import("pyam")

```

```{r define function}
#-------------------------#
#       Functions         #
#-------------------------#

#define functions
##function that calculates CAGR
CAGR_calculator <- function(starting_value, end_value, n ) {
  rate = (end_value / starting_value) ^(1/n)
  rate = rate - 1
  return(rate)
}

##function that calculate linear growth
LG_calculator <- function(starting_value, end_value,n){
  rate =  (end_value / starting_value)^(1/n) - 1
  return(rate)
}

##function that creates df for certain region based on years and growth rate
forecaster <- function(starting_year, starting_value, years, 
                       growth_rate, region, growth_type){
  #create new data frame
  new_df <- data.frame(
    region = character(),
    value = numeric(),
    year = numeric()
  )
  #iterate over years
  for (i in seq(years)){
    n = i
    value = starting_value * ((1+growth_rate)**n)
    #convert data into vector
    new_row <- c(region = region, value = value, year = years[i])
    #append to dataframe while avoid loosing column title
    new_df[nrow(new_df) + 1,] <- new_row
  }
  #return new_df
  return(new_df)
}


#pack running forecast in one function
run_forecast <- function(historic_values, #follow format region, value, year
                         starting_year, #most recent year with data (2020)
                         list_of_growth_rates, #list of growth rates for each region
                         list_of_years, #list of years to run projection for
                         list_of_regions,#list of regions to be included - must be in historic_values
                         scenario_name)#name that scenario should get in csv file
{
  #generate empty list
  df_list <- list()
  ##add starting values
  df_list[[1]] <- historic_values %>% filter(year==starting_year)
  ##add starting index
  index = 1
  for (r in list_of_regions){
    index = index + 1 #iterate index
    #run forecaster function
    new_df = forecaster(starting_year =  starting_year,
                        starting_value = filter(historic_values, region == r, year==starting_year)$value,
                        years = list_of_years,
                        growth_rate = filter(growth_rates,region == r)$gr,
                        region = r)
    #add to df_list
    df_list[[index]] <- new_df
  }
  #combine df_list
  scen_data = do.call(rbind, df_list)
  #name scenario
  scen_data <- scen_data %>%
    #convert value column to numeric
    mutate(value = as.numeric(value)) %>%
    #add column to identify scenario
    mutate(scenario = scenario_name)%>%
    #add identifier at beginning of df
    relocate(scenario, .before = 1)
  return(scen_data)
}
```

```{r set global variables}
#-------------------------#
#    Global Variables     #
#-------------------------#

#create dataframe that will be used to export scenarios in csv formate
export_df <- data.frame()

#create list with relevant regions
regions_filter = c(
  "North America; primarily the United States of America and Canada",
  "Eastern and Western Europe (i.e., the EU28)",
  "Countries of Latin America and the Caribbean",
  "Countries of South Asia; primarily India",
  "Countries of centrally-planned Asia; primarily China",
  "Pacific OECD",
  "Reforming Economies of Eastern Europe and the Former Soviet Union; primarily Russia",
  "Countries of the Middle East; Iran, Iraq, Israel, Saudi Arabia, Qatar, etc.",
  "Countries of Sub-Saharan Africa"
)
```

```{r transport}
#-------------------------#
#       Transport         #
#-------------------------#
#load data
df = pyam$read_iiasa(
  'ar6-public',
  model='REMIND-MAgPIE 2.1-4.2',
  #scenario='2.1-4.2_SusDev_SDP-PkBudg1000',
  variable=c('Energy Service|Transportation|Passenger'),
  meta=c('category')
)
#get right format
df = df$data

#get starting values
hist_data <- c("2005", "2010", "2015", "2020")
#look at historic values
hist_vals <- df %>%
  #filter on year, scenario and region
  filter((year %in% hist_data) & (scenario == 'SusDev_SDP-PkBudg1000') & (region %in% regions_filter)) %>%
  select(region, value, year)

hist_vals %>%
  ggplot(aes(x=year, y=value, color=region))+
  geom_line()

#specify years for scenarios
years <- c(2025, 2030, 2035, 2040, 2045, 2050)
#get regions for scenarios
regions_list <- unique(hist_vals$region)

####----SCENARIO 1 TRANSPORT aggu----####
#AGGREGATE UTILITARIAN
##set growth rates individually via end value
##set end values in 2050; 0 means no change to present day
#find values via
# hist_vals %>% filter(year == 2020) %>% arrange(desc(value)) %>% mutate(value = value + 1000)
end_val = c(
  10295.520 , #Eastern and Western Europe (i.e., the EU28)
  9572.598 , #North America; primarily the United States of America and Canada
  6456.067 , #Countries of centrally-planned Asia; primarily China
  6042.791 , #Countries of the Middle East; Iran, Iraq, Israel, Saudi Arabia, Qatar, etc.
  5329.384 , #Countries of South Asia; primarily India
  5184.672 , #Countries of Latin America and the Caribbean
  3106.586 , #Reforming Economies of Eastern Europe and the Former Soviet Union; primarily Russia
  2554.980 , #Countries of Sub-Saharan Africa
  2396.735 ) #Pacific OECD

#add growth rates
growth_rates <- hist_vals %>%
  filter(year == 2020) %>%
  arrange(desc(value)) %>%
  mutate(end_val = end_val) %>%
  mutate(gr = if_else(end_val == 0,
                      0, 
                      LG_calculator(starting_value = value,
                                      end_value = end_val,
                                      n = length(years))))

#FORECAST
#create empty df
df_list <- list()
##add starting values
df_list[[1]] <- hist_vals %>% filter(year==2020)
##add starting index
index = 1
for (r in regions_list){
  index = index + 1 #iterate index
  new_df = forecaster(starting_year =  2020,
            starting_value = filter(hist_vals, region == r, year==2020)$value,
            years = years,
            growth_rate = filter(growth_rates,region == r)$gr,
            region = r)
  #add to data list
  df_list[[index]] <- new_df
}

scen_data = do.call(rbind, df_list)
#name scenario
scen_data <- scen_data %>%
  #convert value column to numeric
  mutate(value = as.numeric(value)) %>%
  #add column to identify scenario
  mutate(scenario = "transport_agg")%>%
  #add identifier at beginning of df
  relocate(scenario, .before = 1)

scen_data %>%
  ggplot(aes(x=year, y=value, group=region, color=region))+
  geom_line() +
  ggtitle("Transport: Agg utilitarian")+
  ylim(0, 12000)

#ADD TO EXPORT DF
export_df <- rbind(export_df, scen_data)


####----SCENARIO 2 TRANSPORT prio----####
#PRIORITARIAN
##set growth rates individually via end value
##set end values in 2050; 0 means no change to present day
end_val = c(
  0, #Eastern and Western Europe (i.e., the EU28)
  0, #North America; primarily the United States of America and Canada
  0, #Countries of centrally-planned Asia; primarily China
  0, #Countries of the Middle East; Iran, Iraq, Israel, Saudi Arabia, Qatar, etc.
  4800, #Countries of South Asia; primarily India
  4650, #Countries of Latin America and the Caribbean
  3800, #Reforming Economies of Eastern Europe and the Former Soviet Union; primarily Russia
  3500, #Countries of Sub-Saharan Africa
  3400) #Pacific OECD

#add growth rates/ uniform linear growth rates
growth_rates <- hist_vals %>%
  filter(year == 2020) %>%
  arrange(desc(value)) %>%
  mutate(end_val = end_val) %>%
  mutate(gr = if_else(end_val == 0,
                      0, 
                      LG_calculator(starting_value = value,
                                    end_value = end_val,
                                    n = length(years))))

#FORECAST
#create empty df
df_list <- list()
##add starting values
df_list[[1]] <- hist_vals %>% filter(year==2020)
##add starting index
index = 1
for (r in regions_list){
  index = index + 1 #iterate index
  new_df = forecaster(starting_year =  2020,
                      starting_value = filter(hist_vals, region == r, year==2020)$value,
                      years = years,
                      growth_rate = filter(growth_rates,region == r)$gr,
                      region = r)
  #add to data list
  df_list[[index]] <- new_df
}

scen_data = do.call(rbind, df_list)
#name scenario
scen_data <- scen_data %>%
  #convert value column to numeric
  mutate(value = as.numeric(value)) %>%
  #add column to identify scenario
  mutate(scenario = "transport_pri")%>%
  #add identifier at beginning of df
  relocate(scenario, .before = 1)

scen_data %>%
  ggplot(aes(x=year, y=value, group=region, color=region))+
  geom_line() +
  ggtitle("Transport: Prioritarian")+
  ylim(0, 12000)

#ADD TO EXPORT DF
export_df <- rbind(export_df, scen_data)



####----SCENARIO 3 TRANSPORT egal----####
#convergence at 8000

#GET GROWTH RATES
##get CAGR growth rates for each region
growth_rates <- hist_vals %>%
  filter(year == 2020) %>%
  mutate(gr = CAGR_calculator(starting_value = value,
                                end_value = 8000,
                                n = length(years)))

#FORECAST
#create empty df
df_list <- list()
##add starting values
df_list[[1]] <- hist_vals %>% filter(year==2020)
##add starting index
index = 1
for (r in regions_list){
  index = index + 1 #iterate index
  new_df = forecaster(starting_year =  2020,
                      starting_value = filter(hist_vals, region == r, year==2020)$value,
                      years = years,
                      growth_rate = filter(growth_rates,region == r)$gr,
                      region = r)
  #add to data list
  df_list[[index]] <- new_df
}

scen_data = do.call(rbind, df_list)
#name scenario
scen_data <- scen_data %>%
  #convert value column to numeric
  mutate(value = as.numeric(value)) %>%
  #add column to identify scenario
  mutate(scenario = "transport_ega")%>%
  #add identifier at beginning of df
  relocate(scenario, .before = 1)


scen_data %>%
  ggplot(aes(x=year, y=value, group=region, color=region))+
  geom_line() +
  ggtitle("Transport: Convergence at 8.000 pkm")

#ADD TO EXPORT DF
export_df <- rbind(export_df, scen_data)

####----SCENARIO 4 TRANSPORT suf----####
#countries above threshold stay constant, lowest reach threshold by 2040
#GET GROWTH RATES
##set growth rates individually via end value
##set end values in 2050; 0 means no change to present day
##set threshold
threshold = 2500
end_val = c(
  0, #Eastern and Western Europe (i.e., the EU28)
  0, #North America; primarily the United States of America and Canada
  0, #Countries of centrally-planned Asia; primarily China
  0, #Countries of the Middle East; Iran, Iraq, Israel, Saudi Arabia, Qatar, etc.
  0, #Countries of South Asia; primarily India
  0, #Countries of Latin America and the Caribbean
  threshold, #Reforming Economies of Eastern Europe and the Former Soviet Union; primarily Russia
  threshold, #Countries of Sub-Saharan Africa
  threshold) #Pacific OECD

#add growth rates
growth_rates <- hist_vals %>%
  filter(year == 2020) %>%
  arrange(desc(value)) %>%
  mutate(end_val = end_val) %>%
  mutate(gr = if_else(end_val == 0,
                               0, 
                               CAGR_calculator(starting_value = value,
                                               end_value = end_val,
                                               n = length(years)-2)))

#FORECAST
#create empty df
df_list <- list()
##add starting values
df_list[[1]] <- hist_vals %>% filter(year==2020)
##add starting index
index = 1
for (r in regions_list){
  index = index + 1 #iterate index
  new_df = forecaster(starting_year =  2020,
                      starting_value = filter(hist_vals, region == r, year==2020)$value,
                      years = years,
                      growth_rate = filter(growth_rates,region == r)$gr,
                      region = r)
  #change 2045 and 2050 values to threshold for affected years
  if (new_df$value[6] != new_df$value[5]){ #affected years grow until end
    new_df <- new_df %>%
      mutate(value = if_else(year %in% c(2045, 2050), as.numeric(threshold), as.numeric(value)))
    
  }
  #add to data list
  df_list[[index]] <- new_df
}

scen_data = do.call(rbind, df_list)
#name scenario
scen_data <- scen_data %>%
  #convert value column to numeric
  mutate(value = as.numeric(value)) %>%
  #add column to identify scenario
  mutate(scenario = "transport_suf")%>%
  #add identifier at beginning of df
  relocate(scenario, .before = 1)


scen_data %>%
  ggplot(aes(x=year, y=value, group=region, color=region))+
  geom_line()+
  ggtitle("Transport: Sufficitarian")+
  ylim(0, 12000)

#ADD TO EXPORT DF
export_df <- rbind(export_df, scen_data)


####----SCENARIO 5 TRANSPORT lim----####
#countries below threshold stay constant, highest reach threshold by 2040
#GET GROWTH RATES
##set growth rates individually via end value
##set end values in 2050; 0 means no change to present day
##set threshold
threshold = 7500
end_val = c(
  threshold, #Eastern and Western Europe (i.e., the EU28)
  threshold, #North America; primarily the United States of America and Canada
  0, #Countries of centrally-planned Asia; primarily China
  0, #Countries of the Middle East; Iran, Iraq, Israel, Saudi Arabia, Qatar, etc.
  0, #Countries of South Asia; primarily India
  0, #Countries of Latin America and the Caribbean
  0, #Reforming Economies of Eastern Europe and the Former Soviet Union; primarily Russia
  0, #Countries of Sub-Saharan Africa
  0) #Pacific OECD

#add growth rates
growth_rates <- hist_vals %>%
  filter(year == 2020) %>%
  arrange(desc(value)) %>%
  mutate(end_val = end_val) %>%
  mutate(gr = if_else(end_val == 0,
                      0, 
                      CAGR_calculator(starting_value = value,
                                      end_value = end_val,
                                      n = length(years)-2)))

#FORECAST
#create empty df
df_list <- list()
##add starting values
df_list[[1]] <- hist_vals %>% filter(year==2020)
##add starting index
index = 1
for (r in regions_list){
  index = index + 1 #iterate index
  new_df = forecaster(starting_year =  2020,
                      starting_value = filter(hist_vals, region == r, year==2020)$value,
                      years = years,
                      growth_rate = filter(growth_rates,region == r)$gr,
                      region = r)
  #change 2045 and 2050 values to threshold for affected years
  if (new_df$value[6] != new_df$value[5]){ #affected years grow until end
    new_df <- new_df %>%
      mutate(value = if_else(year %in% c(2045, 2050), as.numeric(threshold), as.numeric(value)))
    
  }
  #add to data list
  df_list[[index]] <- new_df
}

scen_data = do.call(rbind, df_list)
#name scenario
scen_data <- scen_data %>%
  #convert value column to numeric
  mutate(value = as.numeric(value)) %>%
  #add column to identify scenario
  mutate(scenario = "transport_lim")%>%
  #add identifier at beginning of df
  relocate(scenario, .before = 1)


scen_data %>%
  ggplot(aes(x=year, y=value, group=region, color=region))+
  geom_line()+
  ggtitle("Transport: Limitarian")+
  ylim(0, 12000)

#ADD TO EXPORT DF
export_df <- rbind(export_df, scen_data)

```

